Flowchart of Current Algorithm

### 2008-08-15 docs/flowchart/paper.tex
no reason to include paper III algorithm anymore since it's updated in lowMach3/paper.tex
Added the equations that the algorithm section references

\newcommand{\uadvone}{\widetilde{\mathbf{U}}^{\mathrm{ADV},(1)}}
\newcommand{\uadvtwo}{\widetilde{\mathbf{U}}^{\mathrm{ADV},(2)}}
\newcommand{\U}{\mathbf{U}}
\newcommand{\V}{\mathbf{V}}
\newcommand{\uadvstar}{\widetilde{\mathbf{U}}^{\mathrm{ADV,\star}}}
\newcommand{\uadvstarone}{\widetilde{\mathbf{U}}^{\mathrm{ADV,(1),\star}}}
\newcommand{\uadvstartwo}{\widetilde{\mathbf{U}}^{\mathrm{ADV,(2),\star}}}
\newcommand{\wadv}{w^{\mathrm{ADV}}}
\newcommand{\dt}{\Delta t}
\newcommand{\dr}{\Delta r}
\newcommand{\dx}{\Delta \x}
\newcommand{\dy}{\Delta r}
\newcommand{\etal}{\it et al.}

\newcommand{\etarho}{\eta_{\rho}}
\newcommand{\etaXk}{\eta_{\rho X_k}}
\newcommand{\etarhoh}{\eta_{\rho h}}

\newcommand{\ubold}{\mathbf{U}}
\newcommand{\ut}{\widetilde{\ubold}}
\newcommand{\wt}{\tilde{w}}

\newcommand{\hp}{\widetilde{h}}

\newcommand{\omegadot}{\dot{\omega}}
\newcommand{\er}{\mathbf{e}_r}
\newcommand{\nx}{\mathbf{n}_x}

\newcommand{\rhobar}{\overline{\rho}}

\newcommand{\x}{\mathbf{x}}
\newcommand{\rp}{r^\prime}
\newcommand{\rhop}{\rho^\prime}

\newcommand{\gcc}{\mathrm{g~cm^{-3} }}

\newcommand{\ext}{\mathrm{ext}}
\newcommand{\eff}{\mathrm{eff}}

\title{Flowchart}

@@

@@

@@

\item {\bf React State}$(\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp})\rightarrow$
$(\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp})$ \newline
is the process by which we evolve the species and enthalpy from $X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following system of equations 
over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )\enskip.
\end{eqnarray}
using the stiff ordinary differential equation
 integration methods provided by the {\tt VODE} package (cite VODE).
The absolute error tolerances are set to $10^{-12}$ for the species,
 and a relative tolerance of $10^{-5}$ is used for the temperature.  
The integration yields the new values of the mass fractions, $X_k^{\outp}$.  
Equation (ref{eq:tempreactC}) is derived from equation (ref{eq:tempenthalpy}) by assuming that the pressure is constant during the burn state.  
In evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  
In theory, this means evaluating the equation of state for each right-hand side evaluation that {\tt VODE} requires.  
In practice, we freeze $c_p$ and $\xi_k$ at the start of the integration time step and compute them using 
$\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs to the equation of state.  
We note that while temperature is evolved when solving these equations, we do not keep the final temperature, 
nor do we use it to compute the final change in enthalpy.  
Therefore, $T^{\outp} = T^{\inp}$.  Also, note that the density remains unchanged during 
the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.

@@

\item {\bf Advect Base}$(\rho_0^{\inp}, p_0^{\inp}, \beta_0^{\inp}, w_0^{\inp}, \psi^{\inp} ) \rightarrow 
(\rho_0^{\outp}, \rho_0^{\outp,n+\myhalf}, p_0^{\outp})$ 
\newline
is the process by which we update the base state through $\dt$ in time given the
radial velocity $w_0^{\inp}.$ Here we discuss
the algorithm for plane-parallel geometries.  
The base state arrays are all one-dimensional 
in the radial coordinate; we think of the base state quantities as defined
at cell centers, with $w_0$ and $\beta_0$ defined halfway in-between.\\ \\
a)  For the density update, we discretize equation (ref{eq:rho0upd\_new}), neglecting the $\etarho$ term,
to compute the new base state density,
\begin{eqnarray}
\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j+\myhalf} -  
         \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right ] \nonumber  \enskip,
\end{eqnarray}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, ${{\rho}_0}^{\outp,\nph}$, are found via the procedure described in
paper II, Appendix A, and are saved for use later in the overall algorithm.\\ \\
b)  For the pressure update, We discretize equation (ref{eq:p0upd\_new}), 
to compute the new base state pressure,
\begin{eqnarray}
p_{0,j}^\outp &=& p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
+ w_{0,j-\myhalf}^\inp \right)
\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\inp \enskip,
\end{eqnarray}
where the interface states are again found via the procedure described in
paper II, Appendix A.
\\

@@

\item At the beginning of each time step, we need an estimate for the time-centered
source term in the velocity divergence constraint (see eq. [ref{eq:constraint}]).
If this is the first step of the calculation ($n=0$), we set
\[
S^{\half,\star} = \frac{S^0 + S^1}{2} \enskip ,
\]
where $S^1$ is found through the iterative process that initializes the calculation.
Otherwise, following the method used in our small-scale low Mach number algorithm
(Cite Bell04), we extrapolate to the half-time using the source
terms at the previous and current time levels
\begin{equation}
S^{\nph,\star} = S^n + \frac{\Delta t^n}{2} \frac{S^n - S^{n-1}}{\Delta t^{n-1}} \enskip .
\end{equation}

@@

\item Construct $w_0^{\nph,\star}$ by integrating equation (ref{eq:divw0}) using 
the lagged $\psi^{\nmh},$
\begin{equation}
\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S}^{\nph,\star} 
- \frac{1}{\gammabar^{n} p_0^{n}} \psi^{\nmh} \enskip . \nonumber
\end{equation}
%for plane-parallel geometries,  or the analogous equation (see eq. [ref{eq:B7}])
%for spherical configurations.  
For spherical geometries, this equation would be modified.
%In the case of spherical geometries use the current $p_0^n,$
%${\Gamma_1}^n_0 = \Gamma_1(p_0^n, \rho_0^n, X_k^n),$ and $g$ as computed from $\rho_0^n.$

\item Using equation (ref{eq:pizero}), define the scaled pressure gradient
\begin{equation}
-\left ( \frac{1}{\rho_0^n} \frac{\partial \pizeroone}{\partial r} \right ) = 
\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} 
+  w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star} \enskip ,
\end{equation} 
where $w_0^{n,\star}$ and $(\partial w_0 / \partial r)^{n,\star}$ are
\begin{equation}
w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\end{equation}
\begin{equation}
\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\end{equation}
If $n=0$, we use $\dt^{-1} = \dt^0$.

@@

The procedure to construct $\uadvone$ is described in detail in Appendix ref{app:advective\_vel}.
We note that  $\uadvone$ satisfies the discrete versions of equations
(ref{eq:udoterzero}) and (ref{eq:tildeconstraint}), 
in particular,
\begin{equation}
\nablab \cdotb \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
\enskip ,
\end{equation}

@@

%The construction of 
%$ \left[ \nabla \cdotb \left(  (\uadvone+w_0^{\nph,\star} \er)  (\rho X_k)^{(1)} \right) \right]^{\nph}$ \\
%and $ \left[ \nabla \cdotb \left( (\uadvone+w_0^{\nph,\star} \er)  (\rho h)^{(1)}\right)  \right]^{\nph}$
%also follows that described in paper II, Appendix A with $\V = \uadvone+w_0^{\nph,\star} \er.$ 
%We note that the time-centering of the $p_0$ terms in equations (ref{eq:step4_rhoh})
%matches as closely as possible the time-centering of the $\beta_0$ in 
%equation (ref{eq:uadvone_satisfies}).

@@

\item Define an edge-centered $\etarho^{\nph,\star}$ and cell-centered $\psi_j^{\nph,\star}$:
\begin{eqnarray}
 \etarho^{\nph,\star}   &=&  {\rm {\bf Avg}} \left( (\uadvone \cdotb \er + w_0^{\nph,\star}) 
\rho^{(1),\nph,\star} \right) - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} 
 \nonumber \enskip , \\
 \psi_j^{\nph,\star}  &=&  \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph,\star} + \eta_{\rho,j+\myhalf}^{\nph,\star}\right) g \enskip .
\end{eqnarray}
For spherical geometries, $\psi_j^{\nph,\star}$ would have a different form. 
%be computed via equation (ref{eq:spherical_psi_num}).

@@

\item Construct $w_0^{\nph}$ by integrating equation (ref{eq:divw0}),
\begin{equation}
\frac{\partial w_0^{\nph}}{\partial r} =  \overline{S}^{\nph,\star} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}} \psi^{\nph,\star} \enskip , \nonumber
\end{equation}
for plane-parallel geometries, where 
$\gammabar^{\nph,\star} = (\gammabar^{n} + \gammabar^{n+1,\star} ) / 2 $ and
$ p_0^{\nph,\star} = (p_0^{n} + p_0^{n+1,\star} ) / 2.$
%In spherical configurations we use the analogous equation (see eq. [ref{eq:B7}]). 
%In the case of spherical geometries use $p_0^{n+1,\star},$
%${\Gamma_1}^{n+1,\star}_0 = \Gamma_1(p_0^{n+1,\star}, \rho_0^{n+1,\star}, X_k^{n+1,\star}),$ and 
%$g$ as computed from $\rho_0^{n+1,\star}.$

\item Using equation (ref{eq:pizero}), define
\begin{equation}
-\left ( \frac{1}{\rho_0^n} \frac{\partial \pizerotwo}{\partial r} \right ) = 
\frac{w_0^{\nph} - w_0^\nmh}{\myhalf(\dt^n+\dt^{n-1})} 
+ w_0^n \left(\frac{\partial w_0}{\partial r}\right)^n.
\end{equation}
where $w_0^n$ and $(\partial w_0 / \partial r)^{n}$ are defined analogously to 
equations (ref{eq:w0nstar}) and (ref{eq:dw0drnstar}).

@@

The procedure to construct $\uadvtwo$ is described in detail in Appendix ref{app:advective\_vel}
and is analogous to the procedure used in step 2, but with updated values
for $w_0$ and $\pizero.$ We note that  $\uadvtwo$ satisfies the discrete versions of 
equations (ref{eq:udoterzero}) and (ref{eq:tildeconstraint}),
in particular
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) = \beta_0^{\nph,\star} 
\left(S^{\nph} - \overline{S}^{\nph}\right) \enskip .
\end{equation}
where $ \beta_0^{\nph,\star} = ( \beta_0^n +  \beta_0^{n+1,\star}) / 2.$

@@

%The construction of 
%$ \left[ \nabla \cdotb \left(  (\uadvtwo+w_0^{\nph} \er)  (\rho X_k)^{(1)} \right) \right]^{\nph}$ \\
%and $ \left[ \nabla \cdotb \left( (\uadvtwo+w_0^{\nph} \er)  (\rho h)^{(1)}\right)  \right]^{\nph}$
%is similar to that described in paper II, Appendix A with $\V = \uadvtwo+w_0^{\nph} \er.$ 
%For species advection we set $s = X_k,$ then multiply the edge states of $X_k$ by predicted
%edge states of $\rho.$  For the enthalpy update we predict temperature to the edges, convert
%$T$ to $h$ using the EOS, then multiply by $\rho.$
%We note that the time-centering of the $p_0$ terms in Equations (ref{eq:step8_rhoh})
%matches as closely as possible the time-centering of the $\beta_0$ in 
%Equation (ref{eq:uadvtwo_satisfies}).

@@

\item Define an edge-centered $ \etarho^{\nph}$ and a cell-centered $\psi_j^{\nph}$,
\begin{eqnarray}
 \etarho^{\nph}   &=&  {\rm {\bf Avg}} \left( (\uadvtwo \cdotb \er + w_0^{\nph}) 
\rho^{(1),\nph} \right) - w_0^{\nph} \rho_0^{\nph,{\rm pred}} 
 \nonumber \enskip , \\
 \psi_j^{\nph}  &=&  \frac{1}{2} \left(\eta_{rho,j-\myhalf}^{\nph} + \eta_{\rho,j+\myhalf}^{\nph}\right) g \enskip .
\end{eqnarray}
%For spherical geometries, $\psi_j^{\nph}$ would be computed via equation (ref{eq:spherical_psi_num}).
\end{enumerate}

@@

The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
We update the velocity field $\ut^n$ to $\ut^{n+1,\star}$ by discretizing
equation (ref{eq:utildeupd}), 
\begin{eqnarray}
\ut^{n+1,\star} &=& \ut^n - \dt \;
 \left[\left((\uadvtwo+ w_0^{\nph} \er) \cdotb \nablab\right) \ut \right]^\nph
      - \dt \; \left(\uadvtwo \cdotb \er\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \er \nonumber \\
   &&   + \dt \left[ - \frac{1}{\rho^\nph} {\mathbf{G}} \pi^\nmh
        + \frac{1}{\rho_0} {\mathbf{G}} \pizerotwo
        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g \er \right] \enskip ,
\end{eqnarray}
where $\rho^\nph = ( \rho^n + \rho^{n+1} )/2$
and $\mathbf{G}$ approximates a cell-centered gradient from nodal data. 
The construction of $\left[((\uadvtwo+ w_0^{\nph} \er) \cdotb \nablab) \ut \right]^\nph$
is described in paper II, Appendix A, with $\V = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
component of $\ut^n$ individually.

>>>>

\newcommand{\uadvone}{\widetilde{\mathbf{U}}^{\mathrm{ADV},(1)}}
\newcommand{\uadvtwo}{\widetilde{\mathbf{U}}^{\mathrm{ADV},(2)}}
\newcommand{\V}{\mathbf{V}}
\newcommand{\dt}{\Delta t}
\newcommand{\dr}{\Delta r}

\newcommand{\etarho}{\eta_{\rho}}

\newcommand{\ubold}{\mathbf{U}}
\newcommand{\ut}{\widetilde{\ubold}}

\newcommand{\omegadot}{\dot{\omega}}
\newcommand{\er}{\mathbf{e}_r}

@@

\title{Flowchart of Current Algorithm}

@@

\section{Equations}
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \label{eq:VODE1C} \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )\enskip. \label{eq:tempreactC}
\end{eqnarray}
\begin{equation}
\label{eq:tempenthalpy}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
 - \sum_k \rho \xi_k {\omegadot}_k 
 + \rho \Hnuc + \rho \Hext \right] \enskip , 
\end{equation}
\begin{eqnarray}
\frac{\partial \rhozero}{\partial t} &=& - \nablab \cdotb \left( \rhozero w_0 \er \right)
- \nablab \cdotb \left( \etarho \er \right) \enskip ,
\label{eq:rho0upd_new}
\end{eqnarray}
\begin{equation}
\frac{\partial p_0}{\partial t} = w_0 \rho_0 g  + \etarho g
                                 = - w_0 \frac{\partial p_0}{\partial r} + \etarho g \enskip .
\label{eq:p0upd_new} 
\end{equation}
\begin{equation}
\label{eq:constraint}
\nablab \cdotb \ubold + \alpha \left( \frac{\partial p_0}{\partial t} + \ubold \cdotb \nablab p_0 \right) =
 -\sigma  \sum_k \xi_k \omegadot_k  + 
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k + \sigma \Hnuc + \sigma \Hext \equiv S \enskip ,
\end{equation}
\begin{eqnarray}
\frac{\partial w_0}{\partial r} &=& \frac{1}{\beta_0} \nablab \cdotb (\beta_0 w_0 \er) 
                                    - w_0 \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}  \nonumber \\
                                &=&   \left( \Sbar - \frac{1}{\gammabar p_0}
                                      \frac{\partial p_0}{\partial t} \right)
                                     - w_0 \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}  \nonumber \\
                                &=&   \Sbar - \frac{1}{\gammabar p_0}
                                      \left( \frac{\partial p_0}{\partial t}
                                       + w_0 \frac{\partial p_0}{\partial r}  \right) \nonumber \\
                                &=&   \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip \label{eq:divw0} ,
\end{eqnarray}
\begin{equation}
- \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
\end{equation}
\begin{equation}
\int_{\Omega_H} \ut \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
\end{equation}
\begin{equation}
\nablab \cdotb (\beta_0 \ut )  = \beta_0 \left(S - \Sbar \right )
 \enskip .  \label{eq:tildeconstraint}
\end{equation}
\begin{equation}
w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\label{eq:w0nstar}
\end{equation}
\begin{equation}
\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\label{eq:dw0drnstar}
\end{equation}
\begin{eqnarray}
\frac{\partial w_0}{\partial t} &=& - w_0 \frac{\partial w_0}{\partial r} - 
\frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \label{eq:w0upd} \enskip , \\
\frac{\partial\ut}{\partial t} &=& - \ut \cdotb \nablab \ut - w_0 \frac{\partial \ut}{\partial r}
                                 - \left(\ut \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
                                 - \frac{1}{\rho} \nablab\pi
                                 + \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
                                 - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip ,
\end{eqnarray}
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
\label{eq:nodal_solve}
\end{equation}
\begin{equation}
  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext \enskip .
\label{eq:defineS} 
\end{equation}

@@

\item {\bf React State}$(\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp})\rightarrow$
$(\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp})$ \newline
is the process by which we evolve the species and enthalpy from $X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following system of equations 
over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )\enskip.
\end{eqnarray}
using the stiff ordinary differential equation
 integration methods provided by the {\tt VODE} package (cite VODE).
The absolute error tolerances are set to $10^{-12}$ for the species,
 and a relative tolerance of $10^{-5}$ is used for the temperature.  
The integration yields the new values of the mass fractions, $X_k^{\outp}$.  
Equation (\ref{eq:tempreactC}) is derived from equation (\ref{eq:tempenthalpy}) by assuming that the pressure is constant during the burn state.  
In evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  
In theory, this means evaluating the equation of state for each right-hand side evaluation that {\tt VODE} requires.  
In practice, we freeze $c_p$ and $\xi_k$ at the start of the integration time step and compute them using 
$\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs to the equation of state.  
We note that while temperature is evolved when solving these equations, we do not keep the final temperature, 
nor do we use it to compute the final change in enthalpy.  
Therefore, $T^{\outp} = T^{\inp}$.  Also, note that the density remains unchanged during 
the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.

@@

\item {\bf Advect Base}$(\rho_0^{\inp}, p_0^{\inp}, \beta_0^{\inp}, w_0^{\inp}, \psi^{\inp} ) \rightarrow 
(\rho_0^{\outp}, \rho_0^{\outp,n+\myhalf}, p_0^{\outp})$ 
\newline
is the process by which we update the base state through $\dt$ in time given the
radial velocity $w_0^{\inp}.$ Here we discuss
the algorithm for plane-parallel geometries.  
The base state arrays are all one-dimensional 
in the radial coordinate; we think of the base state quantities as defined
at cell centers, with $w_0$ and $\beta_0$ defined halfway in-between.\\ \\
a)  For the density update, we discretize equation (\ref{eq:rho0upd_new}), neglecting the $\etarho$ term,
to compute the new base state density,
\begin{eqnarray}
\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j+\myhalf} -  
         \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right ] \nonumber  \enskip,
\end{eqnarray}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, ${{\rho}_0}^{\outp,\nph}$, are found via the procedure described in
paper II, Appendix A, and are saved for use later in the overall algorithm.\\ \\
b)  For the pressure update, We discretize equation (\ref{eq:p0upd_new}), 
to compute the new base state pressure,
\begin{eqnarray}
p_{0,j}^\outp &=& p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
+ w_{0,j-\myhalf}^\inp \right)
\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\inp \enskip,
\end{eqnarray}
where the interface states are again found via the procedure described in
paper II, Appendix A.
\\

@@

\item At the beginning of each time step, we need an estimate for the time-centered
source term in the velocity divergence constraint (see eq. [\ref{eq:constraint}]).
If this is the first step of the calculation ($n=0$), we set
\[
S^{\half,\star} = \frac{S^0 + S^1}{2} \enskip ,
\]
where $S^1$ is found through the iterative process that initializes the calculation.
Otherwise, following the method used in our small-scale low Mach number algorithm
(Cite Bell04), we extrapolate to the half-time using the source
terms at the previous and current time levels
\begin{equation}
S^{\nph,\star} = S^n + \frac{\Delta t^n}{2} \frac{S^n - S^{n-1}}{\Delta t^{n-1}} \enskip .
\end{equation}

@@

\item Construct $w_0^{\nph,\star}$ by integrating equation (\ref{eq:divw0}) using 
the lagged $\psi^{\nmh},$
\begin{equation}
\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S}^{\nph,\star} 
- \frac{1}{\gammabar^{n} p_0^{n}} \psi^{\nmh} \enskip . \nonumber
\end{equation}
For spherical geometries, this equation would be modified.

\item Using equation (\ref{eq:pizero}), define the scaled pressure gradient
\begin{equation}
-\left ( \frac{1}{\rho_0^n} \frac{\partial \pizeroone}{\partial r} \right ) = 
\frac{w_0^{\nph,\star} - w_0^\nmh}{(\dt^n+\dt^{n-1})/2} 
+  w_0^{n,\star} \left(\frac{\partial w_0}{\partial r}\right)^{n,\star} \enskip ,
\end{equation} 
where $w_0^{n,\star}$ and $(\partial w_0 / \partial r)^{n,\star}$ are
\begin{equation}
w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\end{equation}
\begin{equation}
\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\end{equation}
If $n=0$, we use $\dt^{-1} = \dt^0$.

@@

The procedure to construct $\uadvone$ is described in detail in Appendix B of paper III.
We note that  $\uadvone$ satisfies the discrete versions of equations
(\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}), 
in particular,
\begin{equation}
\nablab \cdotb \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
\enskip ,
\end{equation}

@@

@@

\item Define an edge-centered $\etarho^{\nph,\star}$ and cell-centered $\psi_j^{\nph,\star}$:
\begin{eqnarray}
 \etarho^{\nph,\star}   &=&  {\rm {\bf Avg}} \left( (\uadvone \cdotb \er + w_0^{\nph,\star}) 
\rho^{(1),\nph,\star} \right) - w_0^{\nph,\star} \rho_0^{\nph,\star,{\rm pred}} 
 \nonumber \enskip , \\
 \psi_j^{\nph,\star}  &=&  \frac{1}{2} \left(\eta_{\rho,j-\myhalf}^{\nph,\star} + \eta_{\rho,j+\myhalf}^{\nph,\star}\right) g \enskip .
\end{eqnarray}
For spherical geometries, $\psi_j^{\nph,\star}$ would have a different form.

@@

\item Construct $w_0^{\nph}$ by integrating equation (\ref{eq:divw0}),
\begin{equation}
\frac{\partial w_0^{\nph}}{\partial r} =  \overline{S}^{\nph,\star} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}} \psi^{\nph,\star} \enskip , \nonumber
\end{equation}
for plane-parallel geometries, where 
$\gammabar^{\nph,\star} = (\gammabar^{n} + \gammabar^{n+1,\star} ) / 2 $ and
$ p_0^{\nph,\star} = (p_0^{n} + p_0^{n+1,\star} ) / 2.$

\item Using equation (\ref{eq:pizero}), define
\begin{equation}
-\left ( \frac{1}{\rho_0^n} \frac{\partial \pizerotwo}{\partial r} \right ) = 
\frac{w_0^{\nph} - w_0^\nmh}{\myhalf(\dt^n+\dt^{n-1})} 
+ w_0^n \left(\frac{\partial w_0}{\partial r}\right)^n.
\end{equation}
where $w_0^n$ and $(\partial w_0 / \partial r)^{n}$ are defined analogously to 
equations (\ref{eq:w0nstar}) and (\ref{eq:dw0drnstar}).

@@

The procedure to construct $\uadvtwo$ is described in detail in Appendix B of paper III
and is analogous to the procedure used in step 2, but with updated values
for $w_0$ and $\pizero.$ We note that  $\uadvtwo$ satisfies the discrete versions of 
equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}),
in particular
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) = \beta_0^{\nph,\star} 
\left(S^{\nph} - \overline{S}^{\nph}\right) \enskip .
\end{equation}
where $ \beta_0^{\nph,\star} = ( \beta_0^n +  \beta_0^{n+1,\star}) / 2.$

@@

@@

\item Define an edge-centered $ \etarho^{\nph}$ and a cell-centered $\psi_j^{\nph}$,
\begin{eqnarray}
 \etarho^{\nph}   &=&  {\rm {\bf Avg}} \left( (\uadvtwo \cdotb \er + w_0^{\nph}) 
\rho^{(1),\nph} \right) - w_0^{\nph} \rho_0^{\nph,{\rm pred}} 
 \nonumber \enskip , \\
 \psi_j^{\nph}  &=&  \frac{1}{2} \left(\eta_{rho,j-\myhalf}^{\nph} + \eta_{\rho,j+\myhalf}^{\nph}\right) g \enskip .
\end{eqnarray}
\end{enumerate}

@@

The velocity update happens analogously to paper II, using $S^{n+1}$ from step 10.
We update the velocity field $\ut^n$ to $\ut^{n+1,\star}$ by discretizing
equation (\ref{eq:utildeupd}), 
\begin{eqnarray}
\ut^{n+1,\star} &=& \ut^n - \dt \;
 \left[\left((\uadvtwo+ w_0^{\nph} \er) \cdotb \nablab\right) \ut \right]^\nph
      - \dt \; \left(\uadvtwo \cdotb \er\right)  \left(\frac{\partial w_0^{\nph}}{\partial r} \right) \er \nonumber \\
   &&   + \dt \left[ - \frac{1}{\rho^\nph} {\mathbf{G}} \pi^\nmh
        + \frac{1}{\rho_0} {\mathbf{G}} \pizerotwo
        - \frac{(\rho^\nph-\rhozero^\nph)}{\rho^\nph} g \er \right] \enskip ,
\end{eqnarray}
where $\rho^\nph = ( \rho^n + \rho^{n+1} )/2$
and $\mathbf{G}$ approximates a cell-centered gradient from nodal data. 
The construction of $\left[((\uadvtwo+ w_0^{\nph} \er) \cdotb \nablab) \ut \right]^\nph$
is described in paper II, Appendix A, with $\V = \uadvtwo+w_0^{\nph} \er$ and~$s$ set to each
component of $\ut^n$ individually.

Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}),
\[
\nablab \cdotb \left(\beta_0^{\nph} \ut^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)\enskip,
\]
by defining ${\bf V} = \ut^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ and solving
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
\end{equation}
for nodal values of $\phi$, where
$\beta_0^\nph = ( \beta_0^n + \beta_0^{n+1} )/2$ and
$L_\beta^\rho$ is the standard bilinear
finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
In this step, $D$ is a discrete second-order operator that approximates the 
divergence at nodes from cell-centered data and satisfies
$\mathbf{G} = -D^T.$ 
(See (cite almgrenBellSzymczak:1996) for a detailed discussion of this
approximate projection; see (cite almgren:bell:crutchfield) for a discussion
of this particular form of the projection operand.)  
We solve the linear system of equation (\ref{eq:nodal_solve}) using multigrid V-cycles with Gauss-Seidel relaxation.
